Non-word repetition in Yélî Dnye

### 2020-09-14 manuscript.tex
addressed middys comments

\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[english,,man,floatsintext]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provides euro and other symbols
\else % if luatex or xelatex
  \usepackage{unicode-math}
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\usepackage{hyperref}
\hypersetup{
            pdftitle={Non-word repetition in Yélî Dnye},
            pdfkeywords={phonology, non-word repetition, development},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}

@@

\setmainfont{Doulos SIL} % Set main font to Doulos SIL
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

% Table formatting
\usepackage{longtable}
\usepackage{lscape}
% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
\usepackage{multirow}		% Table styling
\usepackage{tabularx}		% Control Column width
\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
\usepackage{threeparttablex}            % Lets threeparttable work with longtable

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\begin{center}\begin{threeparttable}}
%   {\end{threeparttable}\end{center}\end{landscape}}
\newenvironment{lltable}{\begin{landscape}\begin{center}\begin{ThreePartTable}}{\end{ThreePartTable}\end{center}\end{landscape}}

% Enables adjusting longtable caption width to table width
% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

% \setlength{\parindent}{0.5in}
% \setlength{\parskip}{0pt plus 0pt minus 0pt}

% \usepackage{etoolbox}
\makeatletter
\patchcmd{\HyOrg@maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\makeatother
\shorttitle{NWR in Yélî Dnye}
\author{Alejandrina Cristia\textsuperscript{1}\ \& Marisa Casillas\textsuperscript{2}}
\affiliation{
\vspace{0.5cm}
\textsuperscript{1} Laboratoire de Sciences Cognitives et de Psycholinguistique, Département d'Etudes cognitives, ENS, EHESS, CNRS, PSL University\\\textsuperscript{2} Max Planck Institute for Psycholinguistics}
\authornote{All data are made available in a repository in the Open Science Framework. AC acknowledges the support of the Agence Nationale de la Recherche (ANR-17-CE28-0007 LangAge, ANR-16-DATA-0004, ANR-14-CE30-0003, ANR-17-EURE-0017); and the J. S. McDonnell Foundation Understanding Human Cognition Scholar Award.

Correspondence concerning this article should be addressed to Alejandrina Cristia, 29, rue d’Ulm, 75005 Paris, France. E-mail: alecristia@gmail.com}
\keywords{phonology, non-word repetition, development\newline\indent Word count: ~xxx words}
\usepackage{lineno}

\linenumbers
\usepackage{csquotes}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[shorthands=off,main=english]{babel}
\else
  % load polyglossia as late as possible as it *could* call bidi if RTL lang (e.g. Hebrew or Arabic)
  \usepackage{polyglossia}
  \setmainlanguage[]{english}
\fi

\title{Non-word repetition in Yélî Dnye}

\date{}

\abstract{
In nonword repetition (NWR) studies, participants are presented auditorily with an item that is phonologically legal but lexically meaningless in the local language. Accuracy is thought to reflect long-term phonological knowledge as well as online phonological working memory and flexible production patterns.In this study, we report on NWR results among children learning Yêly Dnyé, an isolate spoken in Rossel Island, PNG, with an unusually dense phonological inventory. This study contributes to three lines of research. First, we document that non-word items containing typologically rare sounds are repeated accurately less often that non-words containing more common sounds. Second, we document rather weak effects of item length, contributing to other research suggesting that length effects may be language-specific. Third, we do not find strong individual variation effects in this population, contrary to previous results documenting strong age-related effects. Together, these data provide a unique view of online phonological processing in a seldom-studied language, and contribute to both typological and language acquisition research.
}

@@

\begin{itemize}
\tightlist
\item
  look through this manuscript for bits of text to be written by \textbf{mc}
\item
  help with table 1 (formatting \& making it appear, because it's broken by the phonological representations)
\item
  do you think Figure 2 should be done with phonetic symbols? If yes, could you try to implement that?
\item
  double check demo -- probably start with the xlsx version that has the children's first names, date of birth, and date of test. I looked again at my local files (still haven't checked the labs' ones) and I'm surprised I have maternal education for more children than I have birth order for\ldots{} If you correct this, to generate the csv I also round the age, assign age.rounded to reported age (in years) when DOB is missing, remove first names, DOB, DOTest and age.precise before pushing to github
\item
  double check my frequency entries in segments.xlsx using the search function in \url{https://phoible.org/parameters}; take a look at \url{http://phoible.github.io/conventions/} in case you see something about double articulation (I didn't find tp or lBj)
\item
  double check phonological implementation of stimuli in stimuli.xlsx
\item
  for both of the above, if you make any changes, they need to be re-saved as UTF16 txt file, and then converted (see remember above)
\item
  double check also stim\_seg\_freq -- why are some items still with NA freq?
\item
  consent, \enquote{correct}, \enquote{accuracy}, \enquote{performance}
\item
  draft discussion
\end{itemize}

\hypertarget{todo-later}{%
\subsection{TODO later}\label{todo-later}}

\begin{itemize}
\tightlist
\item
  probably turn summary of types of error into a table; use proportion of correctly produced rather than number; use first attempts rather than the whole data set DONE
\item
  maybe add to the last ana the prevalence (ie out of all the gh which proportion get transformed) DONE I think\ldots{} (this item looks the same as one of the three above)
\item
  maybe add analyses of score as a function of frequency (phones, diphones) in MC's corpus? (mc suggested, but I don't think we need it\ldots{}) - waiting until we talk about this
\end{itemize}

@@

NWR has been used to seek answers to a variety of theoretical questions, including what the links between phonology, working memory, and the lexicon are (Bowey, 2001), and to assess the extent to which apparent phonological constraints found in the lexicon affect online production (Gallagher, 2014). It is also frequently used as for applications, notably as a diagnostic for language delays and disorders (Estes, Evans, \& Else-Quest, 2007). Since non-words can be generated in any language, it has attracted the attention of researchers working in multilingual and linguistically diverse environments, particularly in Europe (Action, 2009; Meir, Walters, \& Armon-Lotem, 2016).

In this study, we report on NWR results among children learning Yêly Dnyé, an isolate spoken in Rossel Island, PNG, with an unusually dense phonological inventory. The present report uses these unique data to contribute to several lines of research. First, we made sure that some of the non-word items contained typologically rare and/or challenging sounds, so that we can contribute to the study of whether rare sounds are disadvantaged in perception and/or production, both in terms of overall accuracy of repetition and patterns of mispronunciation. Second, we look at the effects of item length, since previous NWR research has uncovered variable effects. Third, we contribute to the basic NWR literature by contributing an additional data point for children learning a rarely-studied language and culture, and further investigating whether there are structured sources that account for individual variation.

@@

Yélî Dnye is an isolate language (presumed Papuan) spoken by approximately 7000 people residing on Rossel Island, which is found at the far end of the Louisiade Archipelago in Milne Bay Province, Papua New Guinea. The Yélî sound system, much like its baroque grammatical system (Levinson, n.d.), is unlike any other in the region.

@@

Most words in Yélî Dnye are bisyllabic (\textasciitilde{}50\%), with monosyllabic words (\textasciitilde{}40\%) appearing most commonly after that, and with tri-and-above syllablic words appearing least frequently (\textasciitilde{}10\%; based on \textgreater{}5800 lexemes in the current Yélî Dnye dictionary, Levinson, n.d.). The vast majority of syllables use a CV format. A small portion of the lexicon features words with a final closed syllable, limited to codas of -/m/, -/p/, or -/j/ (e.g., \enquote{ndap} (Spondylus shell) /ṇṭæp/). However, in spontaneous speech an epenthetic /ɯ/ is often inserted after word-final coda, yielding a CVCV structure instead (e.g., \enquote{ndapî} /\enquote{ṇṭæ.pɯ/). This process is used frequently with English loan words that have a coda (e.g., 'ponî} (phone) /\enquote{pɔ.nɯ/). An even smaller portion of the lexicon features words starting with a vowel (e.g., 'ala} (here) /æ.'læ/), but these are limited to /æ/-. Finally, the lexicon features a handful of single-vowel grammatical morphemes (see Levinson (n.d.) for details).

While all consonants can appear in the first syllable of the word, in bisyllabic roots, the second syllable onset consonant is almost exclusively drawn from a subset of the simplex consonants (in order of descending frequency: /m/, /ṭ/, /n/, /p/, /l/, /w/, /j/, /pʷ/, /ŋ/, /k/, /t/, /ɣ/, /mʷ/). In monosyllabic stems, oral vowels (short: 43\% and long: 31\%) are more common than nasal ones (short: 4\% and long: 22\%), with the first syllable of bisyllabic stems following a similar pattern. Most vowels in the second syllable of bisyllabic stems are also short and oral, with a potential preference for rounding harmony between the first and second vowels. Stress typically falls on the first syllable of bisyllabic words, on the first and third syllables of tetrasyllabic words, and on the first or second syllable of trisyllabic words, depending on its vocalic features (see Levinson (n.d.) for details).

\hypertarget{intro-to-the-people-mc---please-feel-free-to-throw-away-anything-that-is-not-useful}{%
\subsubsection{\texorpdfstring{Intro to the people (\textbf{mc}) - please feel free to throw away anything that is not useful!}{Intro to the people (mc) - please feel free to throw away anything that is not useful!}}\label{intro-to-the-people-mc---please-feel-free-to-throw-away-anything-that-is-not-useful}}

Most speakers of the language grow up speaking it monolingually within their swidden horticulturalist communities until they begin attending school around the age of 7; school instruction is in English. While monolingual Yélî Dnye upbringing is common, multilingual families are not unusual, particularly in the region around the Catholic Mission (1953--present) where there appears to be a higher incidence of married-in mothers from other islands (Brown \& Casillas, n.d.), the same region in which the current data were collected. Children in these multilingual families grow up speaking Yélî Dnye and English, Tok Pisin, and/or some other language(s) from the region.

Little is known about language development in children growing up in Rossel Island, a community of primarily subsistence farmers who tend to reside in close-knitted villages where child care is distributed across many individuals, and who typically speak Yélî Dnyé, a phonologically and lexically complex language.

\begin{itemize}
\tightlist
\item
  usually monolingual at home
\item
  schooling in English but it starts at age XX, so not relevant here
\item
  however, some use of English due to immigrants \& children of immigrants
\item
  children spend a lot of time with other children
\item
  most parents are subsistence farmers
\item
  parental education generally varies between XX and YY
\end{itemize}

@@

There is some variation in the presentation procedure and structure of items found in previous NWR work. For example, items are often presented orally by the experimenter (Torrington Eaton, Newman, Ratner, \& Rowe, 2015), although an increasing number of studies have turned to playing back the stimuli in order to have greater control of the stability of the presentation (Brandeker \& Thordardottir, 2015). Additionally, while some studies have used 10-15 non-words, others have employed up to 46 unique items (Piazzalunga, Previtali, Pozzoli, Scarponi, \& Schindler, 2019). Often, authors modulate structural complexity, typically measured in terms of item length (measured in number of syllables) and/or syllable structure (open as opposed to closed syllables, Gallon, Harris, \& Van der Lely, 2007).

@@

Naturally, designing the task in this way may render the study of individual variaiton within the population more difficult because different children are exposed to different items. However, a review of previous work on individual variation suggested to us that many individual differences effects are relatively small, and would not be detectable with the sample size that we could collect in a given visit.

That said, we contribute to the literature by also reporting descriptive analyses of individual variation that could potentially be integrated in meta-analytic efforts. Based on previous work, we looked at potential improvements with age (Farmani et al., 2018; Kalnak, Peyrard-Janvid, Forssberg, \& Sahlén, 2014; Vance, Stackhouse, \& Wells, 2005), and potential negative effects of bilingual exposure (Brandeker \& Thordardottir, 2015; Meir \& Armon-Lotem, 2017; Meir et al., 2016). Previous work typically finds no significant differences as a function of maternal education (e.g., Farmani et al., 2018; Kalnak et al., 2014; Meir \& Armon-Lotem, 2017) or child gender (Chiat \& Roy, 2007). Although previous research has not often investigated potential effects of birth order on NWR, there is a sizable literature on these effects in other language tasks (Havron et al., 2019), and therefore we report on those too.

@@

In view of the hypothesis-driven nature of this work, we had considered boosting the interpretational value of this evidence by announcing our analysis plans prior to conducting them. However, we realized that even pre-registering an analysis would be equivocal because we do not have enough power to look at all relationships of interest, and often to detect any of the known effects given their variability across studies. To illustrate this, we portray studies in which children's NWR scores were gathered between 4 and 12 years of age, and reported separately for items that are relatively short (1-2 syllables) versus longer items (3-4 syllables) in Figure 1. Notice that the effect of stimulus length is minuscule among Italian children (Piazzalunga et al., 2019), but considerable among Tsimane' children (Cristia, Farabolini, Scaff, Havron, \& Stieglitz, 2020), where a drop of 40 percentage points is observed at all ages. A similar difference in NWR scores for short versus long items was observed among Arabic children (Jaber-Awida, 2018). Even the effect of age is unstable in this sample. Whereas it is quite clear that children's NWR scores increase in the Italian data, age effects are less stable among Tsimane' children. Therefore, all analyses here are descriptive and should be considered exploratory.

@@

A first list of candidate items was generated in 2018 by selecting simple consonants (\enquote{p}, \enquote{t}, \enquote{d}, \enquote{k}, \enquote{m}, \enquote{n}, \enquote{w}, \enquote{y}) and vowels (\enquote{i}, \enquote{o},\enquote{u}, \enquote{a},\enquote{e}) that were combined into consonant-vowel syllables, further sampling the space of 1- to 4-syllable sequences. These candidates were automatically checked against Levinson's 2015 dictionary and removed from consideration if they appeared in the dictionary. The second author presented them orally to three local research assistants, who were asked to repeat them and further say whether they were real words. Any item for which two or more of the assistants reported them having a meaning or some form of association was excluded.

>>>>

\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[english,,man,floatsintext]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provides euro and other symbols
\else % if luatex or xelatex
  \usepackage{unicode-math}
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\usepackage{hyperref}
\hypersetup{
            pdftitle={Non-word repetition in Yélî Dnye},
            pdfauthor={Alejandrina Cristia~\& Marisa Casillas},
            pdfkeywords={phonology, non-word repetition, development},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

\setmainfont{Doulos SIL} % Set main font to Doulos SIL
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[shorthands=off,main=english]{babel}
\else
  % load polyglossia as late as possible as it *could* call bidi if RTL lang (e.g. Hebrew or Arabic)
  \usepackage{polyglossia}
  \setmainlanguage[]{english}
\fi

\title{Non-word repetition in Yélî Dnye}
\author{Alejandrina Cristia\textsuperscript{1}~\& Marisa Casillas\textsuperscript{2}}
\date{}

\authornote{All data are made available in a repository in the Open Science Framework. AC acknowledges the support of the Agence Nationale de la Recherche (ANR-17-CE28-0007 LangAge, ANR-16-DATA-0004, ANR-14-CE30-0003, ANR-17-EURE-0017); and the J. S. McDonnell Foundation Understanding Human Cognition Scholar Award.

Correspondence concerning this article should be addressed to Alejandrina Cristia, 29, rue d'Ulm, 75005 Paris, France. E-mail: \href{mailto:alecristia@gmail.com}{\nolinkurl{alecristia@gmail.com}}}

\abstract{
In nonword repetition (NWR) studies, participants are presented auditorily with an item that is phonologically legal but lexically meaningless in the local language. Accuracy is thought to reflect long-term phonological knowledge as well as online phonological working memory and flexible production patterns. In this study, we report on NWR results among children learning Yêly Dnyé, an isolate spoken in Rossel Island, PNG, with an unusually dense phonological inventory. This study contributes to three lines of research. First, we document that non-word items containing typologically rare sounds are repeated accurately less often that non-words containing more common sounds. Second, we document rather weak effects of item length, contributing to other research suggesting that length effects may be language-specific. Third, we do not find strong individual variation effects in this population, contrary to previous results documenting strong age-related effects. Together, these data provide a unique view of online phonological processing in a seldom-studied language, and contribute to both typological and language acquisition research.

}

@@

\begin{itemize}
\tightlist
\item
  look through this manuscript for bits of text to be written by \textbf{mc}
\item
  do you think Figure 2 should be done with phonetic symbols? If yes, could you try to implement that?
\item
  double check demo -- probably start with the xlsx version that has the children's first names, date of birth, and date of test. I looked again at my local files (still haven't checked the labs' ones) and I'm surprised I have maternal education for more children than I have birth order for\ldots{} If you correct this, to generate the csv I also round the age, assign age.rounded to reported age (in years) when DOB is missing, remove first names, DOB, DOTest and age.precise before pushing to github
\item
  double check my frequency entries in segments.xlsx using the search function in \url{https://phoible.org/parameters}; take a look at \url{http://phoible.github.io/conventions/} in case you see something about double articulation (I didn't find tp or lBj)
\item
  double check phonological implementation of stimuli in stimuli.xlsx
\item
  for both of the above, if you make any changes, they need to be re-saved as UTF16 txt file, and then converted (see remember above)
\item
  double check also stim\_seg\_freq -- why are some items still with NA freq?
\item
  consent, ``correct'', ``accuracy'', ``performance''
\item
  draft discussion
\end{itemize}

@@

NWR has been used to seek answers to a variety of theoretical questions, including what the links between phonology, working memory, and the lexicon are (Bowey, 2001), and to assess the extent to which apparent phonological constraints found in the lexicon affect online production (Gallagher, 2014). It is also frequently used for applications, notably as a diagnostic for language delays and disorders (Estes, Evans, \& Else-Quest, 2007). Since non-words can be generated in any language, it has attracted the attention of researchers working in multilingual and linguistically diverse environments, particularly in Europe (Action, 2009; Meir, Walters, \& Armon-Lotem, 2016). However, it has been seldom used outside of Europe and North America (with exceptions including Gallagher, 2014, and Cristia, Farabolini, Scaff, Havron, and Stieglitz (2020)).

In this study, we report on NWR results among children learning Yêly Dnyé, an isolate spoken in Rossel Island, PNG, with an unusually dense phonological inventory. The present report uses these unique data to contribute to several lines of research. First, we made sure that some of the non-word items contained typologically rare and/or challenging sounds, so that we can contribute to the study of whether rare sounds are disadvantaged in perception and/or production, both in terms of overall accuracy of repetition and patterns of mispronunciation. Second, we look at the effects of item length, since previous NWR research has uncovered variable effects. Third, we contribute to the basic NWR literature by adding a data point for children learning a rarely-studied language and culture, and further investigating whether there are structured sources that account for individual variation.

@@

Yélî Dnye is an isolate language (presumed Papuan) spoken by approximately 7,000 people residing on Rossel Island, which is found at the far end of the Louisiade Archipelago in Milne Bay Province, Papua New Guinea. The Yélî sound system, much like its baroque grammatical system (Levinson, n.d.), is unlike any other in the region.

@@

Most words in Yélî Dnye are bisyllabic (\textasciitilde{}50\%), with monosyllabic words (\textasciitilde{}40\%) appearing most commonly after that, and with tri-and-above syllabic words appearing least frequently (\textasciitilde{}10\%; based on \textgreater{}5800 lexemes in the current Yélî Dnye dictionary, Levinson, n.d.). The vast majority of syllables use a CV format. A small portion of the lexicon features words with a final closed syllable, limited to codas of -/m/, -/p/, or -/j/ (e.g., `ndap' (Spondylus shell) /ṇṭæp/). However, in spontaneous speech an epenthetic /ɯ/ is often inserted after word-final coda, yielding a CVCV structure instead (e.g., `ndapî' /`ṇṭæ.pɯ/). This process is used frequently with English loan words that have a coda (e.g., 'ponî' (phone) /`pɔ.nɯ/). An even smaller portion of the lexicon features words starting with a vowel (e.g., 'ala' (here) /æ.'læ/), but these are limited to /æ/-. Finally, the lexicon features a handful of single-vowel grammatical morphemes (see Levinson (n.d.) for details).

While all consonants can appear in the first syllable of the word, the second syllable onset consonant in bisyllabic roots is almost exclusively drawn from a subset of the simplex consonants (in order of descending frequency: /m/, /ṭ/, /n/, /p/, /l/, /w/, /j/, /pʷ/, /ŋ/, /k/, /t/, /ɣ/, /mʷ/). In monosyllabic stems, oral vowels (short: 43\% and long: 31\%) are more common than nasal ones (short: 4\% and long: 22\%), with the first syllable of bisyllabic stems following a similar pattern. Most vowels in the second syllable of bisyllabic stems are also short and oral, with a potential preference for rounding harmony between the first and second vowels. Stress typically falls on the first syllable of bisyllabic words, on the first and third syllables of tetrasyllabic words, and on the first or second syllable of trisyllabic words, depending on its vocalic features (see Levinson (n.d.) for details).

\hypertarget{the-yuxe9luxee-community}{%
\subsubsection{The Yélî community}\label{the-yuxe9luxee-community}}

The current study also contributes to increasing the cultural diversity of the populations for which NWR data have been collected. As mentioned previously, this method has been used almost exclusively in urban, industrialized populations, with only one exception to our knowledge (Cristia et al., 2020). We therefore provide some background information tol help understand some aspects of our study, including the estimation of children's ages and the relationship between NWR scores and children's demographic characteristics.

Most speakers of the language grow up speaking it monolingually until they begin attending school around the age of 7;
school instruction is in English. While monolingual Yélî Dnye upbringing is common, multilingual families are not unusual, particularly in the region around the Catholic Mission---the same region in which the current data were collected---, where there is a higher incidence of married-in mothers from other islands (Brown \& Casillas, n.d.). Children in these multilingual families grow up speaking Yélî Dnye and English, Tok Pisin, and/or some other language(s) from the region.

Most Yélî people are swidden horticulturalists, raising a variety of tubers (e.g., yams, manioc, taro), coconut, banana, and more in their gardens. In addition to garden foods, Yélî children grow up eating fish and foraging for shellfish and nuts, which provide a regular source of extra protein. The typical household in our dataset includes seven individuals and is situated among a collection of 4 or more other households, with structures often arranged around an open grassy area. These household clusters are organized by patrilocal relation, such that they typically comprise a set of brothers, their wives and children, and their mother and father, with neighboring hamlets also typically related through the patriline.

Laying 250 nautical miles off the coast of mainland PNG and surrounded by barrier reef, transport to and from the island is both infrequent and irregular. There is no electricity system (we use solar panels) or motor vehicles on the island, save dinghys with outboard motors that are primarily reserved for medical transport, when there is deisel available. There is a medical radio connection at the health clinic (at the Catholic mission) via which news and messages are often passed along. There is also a cellular tower on the island, though it is typically not functioning. As such, access to news and outside resources, particularly international connections and larger data transfers, is typically unavailable. Our data collection is therefore limited to the duration of the researchers' on-island visits.

Despite restricted outside contact, formal education is a priority for many Yélî families. A recent study surveying more than 40 parents of young children in this region of the island found that nearly all parents completed 6 or more years of education, with 65\% of mothers and 49\% of fathers completing the full range of locally available education (\textasciitilde{}9 years), with around half of those parents having pursued an additional 2+ years of education on other islands in the region or on mainland PNG (Casillas, Brown, \& Levinson, 2020). While there are a handful of local schools around the island, these may be well out of walking distance for many children (i.e., more than 1 hour on foot or by canoe each day), and so it is very common for households to billet school children during the weekdays for months at a time (or for whole months if the child's home is very far). Combined with practices of collective child guardianship within their close-knit hamlets, the practical outcome for researchers is that adult consent can often come from a combination of aunts, uncles, adult cousins, and grandparents standing in for the child's parents---to this child assent is also culturally pertinent, as child independence is encouraged and respected from toddlerhood onward (Brown \& Casillas, n.d.).\footnote{Note that in a few cases this does mean that precise details such as child date of birth, parent age in years, and parent education was only able to be approximated as the parents were residing on another part of the island.} Finally, as mentioned, the language of schooling in this region is English, and so the majority of teens and adults on the island speak some English---many speak it very fluently.

While work on Yélî language development is growing (e.g., Brown, 2011, 2014; Brown \& Casillas, n.d.; Casillas et al., 2020; Liszkowski, Brown, Callaghan, Takada, \& de Vos, 2012), our knowledge of children's linguistic development is quite limited, and research into their phonological development in particular has only just begun {[}e.g., Bram's thesis; MC's CFH data{]}. What we do know about the early language environment is that, while child-directed speech is relatively infrequent throughout the day (Casillas et al., 2020), shared caregiving practices and a near-universal fondness for social interaction with young children ensure that speech directed to children comes from all types of speakers: women, men, and other children (Bunce et al., n.d.; Casillas et al., 2020). While speech from women predominates in young children's speech environments on Rossel, as it does elsewhere (Bergelson et al., 2019; Bunce et al., n.d.), there is a significant and marked increase in child-directed speech from other children as infants get older (Bunce et al., n.d.; Casillas et al., 2020). This increase in child-directed speech from other children is attributed to the fact that, starting around age 2, children often spend large swaths of the day playing, swimming, and foraging for fruit, nuts, and shellfish in large (\textasciitilde{}10 people) independent and mixed-age child play groups (Brown \& Casillas, n.d.; Casillas et al., 2020).

@@

The basic structure of NWR involves the participant hearing a non-word, like ``bilik'', which they need to repeat back without changing anything that is contrastive in the language. For instance, in English, a response of ``bilig'' or ``pilik'' will be scored as incorrect; a response ``biilik'', where the vowel is lengthened without change of quality would be scored as correct, because English does not have a pure length contrast.

Aside from this central basic structure, there is some variation in the presentation procedure and structure of items found in previous NWR work. For example, items are often presented orally by the experimenter (Torrington Eaton, Newman, Ratner, \& Rowe, 2015), although an increasing number of studies have turned to playing back the stimuli in order to have greater control of the stability of the presentation (Brandeker \& Thordardottir, 2015). Additionally, while some studies have used 10-15 non-words, others have employed up to 46 unique items (Piazzalunga, Previtali, Pozzoli, Scarponi, \& Schindler, 2019). Often, authors modulate structural complexity, typically measured in terms of item length (measured in number of syllables) and/or syllable structure (open as opposed to closed syllables, Gallon, Harris, \& Van der Lely, 2007).

@@

Naturally, designing the task in this way may render the study of individual variation within the population more difficult because different children are exposed to different items. However, a review of previous work on individual variation suggested to us that effects of individual differences are relatively small, and would not be detectable with the sample size that we could collect in a given visit.

That said, we contribute to the literature by also reporting descriptive analyses of individual variation that could potentially be integrated in meta-analytic efforts. Based on previous work, we looked at potential improvements with age (Farmani et al., 2018; Kalnak, Peyrard-Janvid, Forssberg, \& Sahlén, 2014; Vance, Stackhouse, \& Wells, 2005), and potential negative effects of bilingual exposure (Brandeker \& Thordardottir, 2015; Meir \& Armon-Lotem, 2017; Meir et al., 2016). Previous work typically finds no significant differences as a function of maternal education (e.g., Farmani et al., 2018; Kalnak et al., 2014; Meir \& Armon-Lotem, 2017) or child gender (Chiat \& Roy, 2007). Although past research has not often investigated potential effects of birth order on NWR, there is a sizable literature on these effects in other language tasks (Havron et al., 2019), and therefore we report on those too.

@@

In view of the hypothesis-driven nature of this work, we had considered boosting the interpretational value of this evidence by announcing our analysis plans prior to conducting them. However, we realized that even pre-registering an analysis would be equivocal because we do not have enough power to look at all relationships of interest, and often to detect any of the known effects given their variability across studies. To illustrate this, we portray studies in which children's NWR scores were gathered between 4 and 12 years of age, and reported separately for items that are relatively short (1-2 syllables) versus longer items (3-4 syllables) in Figure 1. Notice that the effect of stimulus length is minuscule among Italian children (Piazzalunga et al., 2019), but considerable among Tsimane' children (Cristia et al., 2020), where a drop of 40 percentage points is observed at all ages. A similar difference in NWR scores for short versus long items was observed among Arabic children (Jaber-Awida, 2018). Even the effect of age is unstable in this sample. Whereas it is quite clear that children's NWR scores increase in the Italian data, age effects are less stable among Tsimane' children. Therefore, all analyses here are descriptive and should be considered exploratory.

@@

A first list of candidate items was generated in 2018 by selecting simple consonants (``p'', ``t'', ``d'', ``k'', ``m'', ``n'', ``w'', ``y'') and vowels (``i'', ``o'',``u'', ``a'',``e'') that were combined into consonant-vowel syllables, further sampling the space of 1- to 4-syllable sequences. These candidates were automatically checked against Levinson's 2015 dictionary and removed from consideration if they appeared in the dictionary. The second author presented them orally to three local research assistants, who were asked to repeat them and further say whether they were real words. Any item for which two or more of the assistants reported them having a meaning or some form of association was excluded.

A second list of candidate items was generated in 2019 by selecting complex consonants and systematically crossing them with all the vowels in the Yélî inventory to produce consonant-vowel monosyllables. As before, items were automatically excluded if they appeared in the dictionary. Additionally, since hearing vowel length in monosyllables in isolation is challenging, any item that had a short/long real word neighbor was filtered out. Since the phonology and phonetics of Yélî is still in the process of being described (Levinson, n.d.), there could have been undocumented constraints that rendered items illegal. Therefore, we made sure that the precise consonant-vowel sequence occurred in some real word in the dictionary (i.e., that there was a longer word included the monosyllable as a subsequence). These candidates were presented to one informant, for a final check that they did not mean anything. Together with the 2018 selection, they were recorded using a headset \textbf{XXX }mc** please fill in** and an Olympus \textbf{XXX }mc** please fill in** from the written form presented together with the same item orally (by the second author). The complete recorded list was finally presented to two more informants, who could repeat all the items and who confirmed there were no real words. Even so, there was one monosyllable that was often identified as a real word (intended ``yî'' /yXX/; identified as ``yi'' /yi/, \emph{tree}). This item is removed from analyses below.
